Parsing Postfix log files

### 2009-04-12 doc/logparser-acronyms.tex
Made corrections based on Ariane's comments for chapters 1 and 2. Acronyms are now expanded the first time they're used in each chapter, rather than just the first time they're used in the thesis.

\newglossaryentry{SMTP}{
    type=\acronymtype,
    name={SMTP},
    description={Simple Mail Transfer Protocol},
    first={Simple Mail Transfer Protocol (SMTP)}
}
\newcommand{\SMTPglossaryDescription}[0]{%
    \acronym{SMTP} is the protocol used for transferring mail between the
    sending and receiving \acronym{MTA}\@.  It is a simple, human readable,
    plain text protocol, making it quite simple to test and debug problems
    with it.  A detailed description of SMTP is beyond the scope of this
    thesis: the original protocol definition is in RFC~821~\cite{RFC821},
    updated in RFC~2821~\cite{RFC2821}; an introductory guide can be found
    at
    \urlLastChecked{http://db.glug-bom.org/lug-authors/philip/docs/mail-stuff/smtp-intro.html}{2009/02/28}
    if more information is needed.%
}
\newglossaryentry{SMTP glossary}{
    name={Simple Mail Transfer Protocol},
    description={
        \SMTPglossaryDescription{}
    }
}

>>>>

\newglossaryentry{SMTP}{
    type=\acronymtype,
    name={SMTP},
    description={Simple Mail Transfer Protocol},
    first={Simple Mail Transfer Protocol (SMTP)}
}
\newcommand{\SMTPglossaryDescription}[0]{%
    \acronym{SMTP} is the protocol used for transferring mail between the
    sending and receiving \acronym{MTA}\@.  It is a simple, human readable,
    plain text protocol, making it quite simple to test and debug problems
    with it.  A detailed description of SMTP is beyond the scope of this
    thesis: the original protocol definition is in
    \acronym{RFC}~821~\cite{RFC821}, which was updated in
    \acronym{RFC}~2821~\cite{RFC2821}.%
}
\newglossaryentry{SMTP glossary}{
    name={Simple Mail Transfer Protocol},
    description={
        \SMTPglossaryDescription{}
    }
}

### 2009-04-12 doc/logparser-background.tex
Made corrections based on Ariane's comments for chapters 1 and 2. Acronyms are now expanded the first time they're used in each chapter, rather than just the first time they're used in the thesis.

This section provides background information helpful in understanding the
remainder of this thesis.  It begins with a discussion of the motivation
underlying the project, followed by a brief introduction to \acronym{SMTP},
and finishes with a longer introduction to Postfix, concentrating on
anti-spam restrictions and policy servers.

@@

\SMTPglossaryDescription{} Despite the simplicity of the protocol many
virus and spam sending programs fail to implement it properly, so requiring
strict adherence to the protocol specification is beneficial in protecting
against spam and viruses.\footnote{Originally all mail servers adhered to
the principle of \textit{Be liberal in what you accept, and conservative in
what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but unfortunately
that principle was written in a friendlier time.  Given the deluge of spam
that mail servers are subjected to daily, a more appropriate maxim could
be: \textit{Require strict adherence to relevant \acronyms{RFC}; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  It is neither as friendly nor as
catchy, but it more accurately reflects the current situation.} A typical
\acronym{SMTP} conversation resembles the following (the lines starting
with a three digit number are sent by the server, all other lines are sent
by the client):

@@

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, other programs,
particularly newer programs, were written by competent programmers or send
mail using well written programs (e.g.\ Postfix or Sendmail on Unix hosts,
Microsoft Outlook on Windows hosts).  Traditionally a mail server would
have done its best to deal with deficient clients, with the intention of
accepting as much mail sent to its users as possible, e.g.\ by ignoring the
absence of a HELO command, or by accepting sender or recipient addresses
that were not enclosed in \texttt{<>}.

>>>>

This chapter provides background information helpful in understanding the
remainder of this thesis.  It begins with a discussion of the motivation
underlying the project, followed by a brief introduction to \acronym{SMTP},
and finishes with a longer introduction to Postfix, concentrating on
anti-spam restrictions and policy servers.

@@

\SMTPglossaryDescription{} Despite the simplicity of the protocol, many
virus and spam sending programs fail to implement it properly, so requiring
strict adherence to the protocol specification is beneficial in protecting
against spam and viruses.\footnote{Originally all mail servers adhered to
the principle of \textit{Be liberal in what you accept, and conservative in
what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but unfortunately
that principle was written in a friendlier time.  Given the deluge of spam
that mail servers are subjected to daily, a more appropriate maxim could
be: \textit{Require strict adherence to relevant \acronyms{RFC}; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  It is neither as friendly nor as
catchy, but it more accurately reflects the current situation.} A typical
\acronym{SMTP} conversation resembles the following (the lines starting
with a three digit number are sent by the server, all other lines are sent
by the client):

@@

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, other programs,
particularly newer programs, were written by competent programmers or send
mail using well written programs (e.g.\ Postfix or Sendmail on Unix hosts,
Microsoft Outlook on Windows hosts).  Traditionally a mail server would
have done its best to deal with deficient clients, with the intention of
accepting as much mail sent to its users as possible, e.g.\ by accepting
sender or recipient addresses that were not enclosed in \texttt{<>}, or by
ignoring the absence of a HELO command.  The HELO command is the first
command sent by the client in the \acronym{SMTP} conversation; it is
required by the protocol, but its original purpose was to detect a server
trying to deliver mail to itself, so some implementations omitted it.

### 2009-04-12 doc/logparser-introduction.tex
Made corrections based on Ariane's comments for chapters 1 and 2. Acronyms are now expanded the first time they're used in each chapter, rather than just the first time they're used in the thesis.

Portions of chapters~\ref{introduction} and~\ref{parser
architecture}--\ref{Evaluation} have previously been published at a
conference~\cite{sgai-2008}, and later reprinted in a
journal~\cite{elsevier-2009}.  Publication of the conference paper was
supported by Science Foundation Ireland RFP~05/RF/CMS002.

>>>>

Portions of chapters~\ref{introduction} and~\ref{parser
architecture}--\ref{Evaluation} have previously been published at an
international conference~\cite{sgai-2008}, and later reprinted in a
journal~\cite{elsevier-2009}.  Publication of the conference paper was
supported by Science Foundation Ireland RFP~05/RF/CMS002.

### 2009-04-12 doc/logparser-results.tex
Made corrections based on Ariane's comments for chapters 1 and 2. Acronyms are now expanded the first time they're used in each chapter, rather than just the first time they're used in the thesis.

Verifying by inspection that the parser correctly deals with all 60,721,709
log lines in the \numberOFlogFILES{} log files is infeasible, but verifying
the parsing of a sample from those log lines is a tractable, if extremely
time consuming, task.  A sample of log lines was obtained by randomly
selecting a log file, and then randomly selecting a block of 6000
contiguous log lines from it (0.00988\% of the total number of log lines).
It is important that the log lines are contiguous, so that all log lines
are present for as many of the mail delivery attempts contained in the
block as possible.  This log segment was parsed with all debugging options
enabled, resulting in 167,448 lines of output.\footnote{A mean of 27.908
lines of output per line of input; each connection has 30 debugging lines,
plus 21 debugging lines per result.  Connections which have been cloned
will have the cloned connection in their debugging output, plus another 33
debugging lines.  Those numbers are approximate, and may vary $\pm{}$ 2.
An approximate linear relationship between the number of log lines and
debugging lines is: $33(connections) + 30(accepted~~mails) + 21(results)$.}
All 167,448 lines were examined in conjunction with the log file segment
and a dump of the resulting database, verifying that for each of the log
lines \parsername{} used the correct rule and invoked the correct action,
which in turn produced the correct result, and the correct data was
inserted in the database.  The log file segment produced 4 warnings, 10
mails remaining in the state tables, 1625 mail delivery attempts correctly
entered in the database, zero false positives, and zero false negatives.

>>>>

Verifying by inspection that the parser correctly deals with all 60,721,709
log lines in the \numberOFlogFILES{} log files is infeasible, but verifying
the parsing of a sample from those log lines is a tractable, if extremely
time consuming, task.  A sample of log lines was obtained by randomly
selecting a log file, and then randomly selecting a block of 6000
contiguous log lines from it (0.00988\% of the total number of log lines).
It is important that the log lines are contiguous, so that all log lines
are present for as many of the mail delivery attempts contained in the
block as possible.  This log segment was parsed with all debugging options
enabled, resulting in 167,448 lines of output.\footnote{A mean of 27.908
lines of output per log line; each connection has 30 debugging lines, plus
21 debugging lines per result.  Connections which have been cloned will
have the cloned connection in their debugging output, plus another 33
debugging lines.  Those numbers are approximate, and may vary $\pm{}$ 2.
An approximate linear relationship between the number of log lines and
debugging lines is: $33(connections) + 30(accepted~~mails) + 21(results)$.}
All 167,448 lines were examined in conjunction with the log file segment
and a dump of the resulting database, verifying that for each of the log
lines \parsername{} used the correct rule and invoked the correct action,
which in turn produced the correct result, and the correct data was
inserted in the database.  The log file segment produced 4 warnings, 10
mails remaining in the state tables, 1625 mail delivery attempts correctly
entered in the database, zero false positives, and zero false negatives.
